---
title: "Ch. 18 - Linear Regression Models - Problem 18.1 Crime and Punishment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


The data in linearRegression_crimePunishment.csv contains the murder rate per capita and the
rate of automobile crimes per 100,000 individuals (both on the log scale) in the ten US States that
have changed their legislation on capital punishment since 1960 (in all cases the states abolished capital punishment). We also include a dummy variable (“law”) that is 1 if the state allows capital punishment in that year, and 0 otherwise. The crime data is from [http://www.disastercenter.com](http://www.disastercenter.com).

```{r}
cp_data <- readr::read_csv(here::here("STAN", "data", 
                                      "linearRegression_crimePunishment.csv")) %>% 
  select(-X1) %>% janitor::clean_names() %>% 
  mutate(state_name = factor(state_name),
         state = as.integer(state),
         law = as.integer(law))
```


# Problem 18.1.1. Graph the data and comment on any trends.

```{r}
library(ggplot2)

p <- ggplot(data = cp_data, aes(x = year)) + 
  geom_line(aes(y = murder), color = "red") + 
  geom_line(aes(y = car), color = "green") + 
  geom_line(aes(y = law), color = "blue") +
  facet_grid(rows = vars(state_name))

p
```

There seems to be some association between the
murder rate and auto crimes. In all cases it is difficult to visually discern an impact of the change
in legislation.

# Problem 18.1.2.

A simple model for murder rates is of the form:

$$
murder_{i,t} ∼ N(\alpha + \beta \times penalty_{i,t} + \gamma \times car_{i,t}, \sigma) 
$$
where we assume that the effect of having the death penalty is given by β, which is assumed
to be the same across all states. We include $car_{i,t}$ – a measure of crimes on automobiles, as an
independent variable to proxy for the contemporaneous underlying level of crime. Estimate this
model and hence determine whether the death penalty acts as a deterrent to murder.

```{r}
library(rstan)

options(mc.cores = parallel:: detectCores())
rstan_options(auto_write = TRUE)

data_list <- list(N       = nrow(cp_data),
                  murder  = cp_data$murder,
                  penalty = cp_data$law,
                  car     = cp_data$car)

set.seed = 1

fit <- stan(here::here("STAN", "problem_18_1_linear_regression", "c_p.stan"), 
            data = data_list, 
            iter = 1000, 
            chains = 4, 
            warmup = 500)
```

```{r}
print(fit, c("alpha", "beta_car", "beta_penalty", "sigma", "lp__"))
```

Where we see that we estimate that the imposition of the death penalty on average raises the
murder rate by on average 24%!

# Problem 18.1.3. 

An alternative model allows there to be state-level effects:

$$
murder_{i,t} ∼ N(\alpha_i + \beta_i \times penalty_{i,t} + \gamma_i \times car_{i,t}, \sigma_i)
$$

where we assume that $\alpha_i ∼ N(\bar{\alpha}, \sigma_\alpha)$, $\beta_i ∼ N(\bar{\beta}, \sigma_\beta)$ and $\gamma_i ∼ N(\bar{\gamma}, \sigma_\gamma)$ (we assume fully heterogeneous estimates for $\sigma$). Estimate the above model and compare the results with the homogeneous coefficient model.

```{r}
data_list <- list(K       = cp_data$state_name %>% unique %>% length,
                  N       = nrow(cp_data),
                  murder  = cp_data$murder,
                  penalty = cp_data$law,
                  car     = cp_data$car,
                  state   = cp_data$state)

set.seed = 1

fit_2 <- stan(here::here("STAN", "problem_18_1_linear_regression", "c_p_hyer.stan"), 
              data = data_list, 
              iter = 2000, 
              chains = 4, 
              warmup = 1000)
```

```{r}
print(fit_2, c("alpha", "beta_car", "beta_penalty", "sigma", "alpha_top",
             "sigma_alpha_top", "beta_car_top", "sigma_car_top",
             "beta_penalty_top", "sigma_penalty_top", "alpha_average",
             "beta_car_average", "beta_penalty_average", "lp__"))
```

A mean effect size of a 25% reduction in murder rates although with a much wider range of
effects.

# Problem extra. (Not in main text but wanted to include) 

Another model allows there to be time
trends in the data:

$$
murder_{i,t} ∼ N(\alpha + \delta_i \times t + \beta \times penalty_{i,t} + \gamma \times car_{i,t}, \sigma) 
$$

where $δ_i ∼ N(\bar{\delta}, σ_\delta)$. Again estimate this model and compare the effect size of the death penalty
across the three models.

```{r}
data_list <- list(K       = cp_data$state_name %>% unique %>% length,
                  N       = nrow(cp_data),
                  murder  = cp_data$murder,
                  penalty = cp_data$law,
                  car     = cp_data$car,
                  state   = cp_data$state,
                  t       = cp_data$year - min(cp_data$year + 1))

set.seed = 1

fit_3 <- stan(here::here("STAN", "problem_18_1_linear_regression", "c_p_hyer_2.stan"), 
              data = data_list, 
              # control = list(max_treedepth = 12),
              iter = 2000, 
              chains = 4, 
              warmup = 1000)
```

```{r}
print(fit_3, c("alpha", "delta", "beta_car", "beta_penalty", "sigma",
               "alpha_top", "sigma_alpha_top", 
               "delta_top", "sigma_delta_top",
               "beta_car_top", "sigma_car_top",
               "beta_penalty_top", "sigma_penalty_top", 
               "alpha_average", 
               "delta_average", 
               "beta_car_average", 
               "beta_penalty_average", 
               "lp__"))
```

```{r}
library(loo)

logLikelihood_homogeneous <- extract_log_lik(fit, "logLikelihood")
WAIC_homogeneous <- waic(logLikelihood_homogeneous)

logLikelihood_state_level_no_trend <- extract_log_lik(fit_2, "logLikelihood")
WAIC_state_level_no_trend <- waic(logLikelihood_state_level_no_trend)

logLikelihood_state_level_with_trend <- extract_log_lik(fit_3, "logLikelihood")
WAIC_state_level_with_trend <- waic(logLikelihood_state_level_with_trend)

compare(WAIC_homogeneous, 
        WAIC_state_level_no_trend, 
        WAIC_state_level_with_trend)
loo_compare(WAIC_homogeneous, 
        WAIC_state_level_no_trend, 
        WAIC_state_level_with_trend)
```

```{r}
LOO_homogeneous <- loo(logLikelihood_homogeneous)
LOO_state_level_no_trend <- loo(logLikelihood_state_level_no_trend)
LOO_state_level_with_trend <- loo(logLikelihood_state_level_with_trend)

compare(LOO_homogeneous, 
        LOO_state_level_no_trend, 
        LOO_state_level_with_trend)
loo_compare(LOO_homogeneous, 
            LOO_state_level_no_trend, 
            LOO_state_level_with_trend)

```

# Problem 18.1.5. 

Critically evaluate the best-performing model and hence any conclusions that
can be drawn from this analysis.


# Some plots

[Plotting MCMC draws using the bayesplot package](https://mc-stan.org/bayesplot/articles/plotting-mcmc-draws.html)

```{r}
library(bayesplot)


posterior <- as.array(fit)
dim(posterior)

# dimnames(posterior)
```

## Posterior uncertainty intervals

```{r}
color_scheme_set("red")
mcmc_intervals(posterior, pars = c("alpha","beta_car", "beta_penalty", "sigma"))
```

```{r}
mcmc_areas(
  posterior,
  pars = c("alpha","beta_car", "beta_penalty", "sigma"),
  prob = 0.8, # 80% intervals
  prob_outer = 0.99, # 99%
  point_est = "mean"
)
```

## Univariate marginal posterior distributions

```{r, eval = FALSE}
color_scheme_set("green")
mcmc_hist(posterior, 
          pars = c(mcmc_areas(posterior,
                              pars = c("alpha", "beta_car", "beta_penalty", "sigma"),
                              prob = 0.8, # 80% intervals
                              prob_outer = 0.99, # 99%
                              point_est = "mean")))
```

```{r}
color_scheme_set("blue")
mcmc_hist(posterior, pars = c("alpha", "beta_car", "beta_penalty", "sigma"),
          transformations = list("sigma" = "log"))
```

```{r}
color_scheme_set("brightblue")
mcmc_hist_by_chain(posterior, pars = c("alpha", "beta_car", "beta_penalty", "sigma"))
```

```{r}
color_scheme_set("purple")
mcmc_dens(posterior, pars = c("alpha", "beta_car", "beta_penalty", "sigma"))
```

```{r}
mcmc_dens_overlay(posterior, pars = c("alpha", "beta_car", "beta_penalty", "sigma"))
```

```{r}
color_scheme_set("teal")
mcmc_violin(posterior, pars = c("alpha", "beta_car", "beta_penalty", "sigma"), probs = c(0.1, 0.5, 0.9))
```


# Bivariate plots

```{r}
color_scheme_set("gray")
mcmc_scatter(posterior, pars = c("alpha", "beta_car"),
             size = 1.5, alpha = 0.5)
```

```{r}
# requires hexbin package
if (requireNamespace("hexbin", quietly = TRUE)) {
  mcmc_hex(posterior, pars = c("alpha", "beta_car"))
}
```

```{r}
color_scheme_set("pink")
mcmc_pairs(posterior, pars = c("alpha", "beta_car", "beta_penalty", "sigma"),
           off_diag_args = list(size = 1.5))
```

